\Title

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: conclusion

%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
To test our model, we will quantify its ability to categorize different motions. Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial size and duration similarly to the motion task defined above.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions $
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$.
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right). Note that these accuracy is computed both in the case is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------

@@

\note{try also a quantization of the weights (binary?)}
%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

Let us highlight some innovations in the contributions presented in this paper. First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we have observed to be less efficient. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif. Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen. Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}. Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs. This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified several limits to our model, which we will now detail. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations.  \note{pourrait parler du poster bernstein? de nadafian?}

Another restriction is that this model is purely feed-forward. Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field. However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}. For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}. Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, Notably to explain non-linear mechanisms~\citep{boutin_pooling_2022}. It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions. However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses. While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).  
\note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}

Such recurrent activity would be essential for implementing predictive or anticipatory processes. This is essential in a neural system, as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}. This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}. \note{Say that this could be implemented by properly setting the timing of the label} As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}. Our proposed model using heterogeneous delays provides an alternate and elegant implementation solution to this problem.
%
\subsection{Perspectives}
%%%-----------------------------------------------------------------
\note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image. This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented. Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}. Thus, there is a dependence within the event stream between motion and orientation information. It would be crucial to study this interdependence further. This could be initiated by training the model on a dataset with labels providing local orientation. \note{we can say we tried and that it works well}. Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them. In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.

>>>>

%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
To test our model, we will quantify its ability to categorize different motions. Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial size and duration similarly to the motion task defined above.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions $
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$.
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right). Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------

@@

% \note{try also a quantization of the weights (binary?)}
%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. % \note{to change if we compare to ghosh instead of HOTs if we follow that route} 
The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

Let us highlight some innovations in the contributions presented in this paper. First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we have observed to be less efficient. Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif. Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen. Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}. Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs. This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified several limits to our model, which we will now detail. First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time. We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions. We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme. Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework. Such an architecture could provide promising speed gains for the calculations. % \note{pourrait parler du poster bernstein? de nadafian?}

Another restriction is that this model is purely feed-forward. Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field. However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}. For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}. Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, Notably to explain non-linear mechanisms~\citep{boutin_pooling_2022}. It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions. However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses. While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).  
% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}

Such recurrent activity would be essential for implementing predictive or anticipatory processes. This is essential in a neural system, as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}. This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}. %\note{Say that this could be implemented by properly setting the timing of the label} 
As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, \ldots) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}. Our proposed model using heterogeneous delays provides an alternate and elegant implementation solution to this problem.
%
\subsection{Perspectives}
%%%-----------------------------------------------------------------
% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image. This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented. Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}. Thus, there is a dependence within the event stream between motion and orientation information. It would be crucial to study this interdependence further. This could be initiated by training the model on a dataset with labels providing local orientation. %\note{we can say we tried and that it works well}. 
Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them. In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
